\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2021-11-20 main.Rmd
change to JSS format

---
title: A template for the *arxiv* style
authors:
  - name: H.Sherry Zhang
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  huize.zhang@monash.edu
  - name: Dianne Cook
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    department: Institute of Statistics
    affiliation: University of Natural Resources and Life Sciences
    location: Vienna, Austria
    email:  ursula.laa@boku.ac.at  
  - name: Nicolas Langrené
    department: 34 Village Street, Docklands VIC 3008 Australia
    affiliation: CSIRO Data61 
    location: Melbourne, Australia
    email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    department: Department of Econometrics and Business Statistics 
    affiliation: Monash University 
    location: Melbourne, Australia
    email:  patricia.menendez@monash.edu 
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.path = "figures/",
                      out.width = "100%", 
                      out.height = "30%")
```

@@

cand <- aus_climate %>% 
  select(id, lat, long, elev, name, wmo_id) %>% 
   as_tibble()

@@

```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", fig.align='center'}
knitr::include_graphics(here::here("figures/cubble-operations/cubble-operations.001.png"))
```

>>>>

---
documentclass: jss
author:
  - name: H. Sherry Zhang
    # department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    #location: Melbourne, Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    # department: Department of Econometrics and Business Statistics
    affiliation: 'Monash University \AND'
    # location: Melbourne, Australia
    # email:  dicook@monash.edu
  - name: Ursula Laa
    # department: Institute of Statistics
    affiliation: 'University of Natural Resources and Life Sciences \AND'
    # location: Vienna, Austria
    # email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    # department: 34 Village Street, Docklands VIC 3008 Australia
    affiliation: 'CSIRO Data61 \AND'
    # location: Melbourne, Australia
    # email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    # department: Department of Econometrics and Business Statistics
    affiliation: 'Monash University \AND'
    # location: Melbourne, Australia
    # email:  patricia.menendez@monash.edu
title:
  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [spatio-temporal data,  "\\proglang{R}"]
  plain:     [spatio-temporal data, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
bibliography: references.bib
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.path = "figures/",
                      out.width = "100%", 
                      out.height = "30%")
options(prompt = 'R> ', continue = '+ ')
```

@@

cand <- aus_climate %>% 
  as_tibble() %>% 
  select(id, lat, long, elev, name, wmo_id)

@@

```{r cubble-operations, echo = FALSE, fig.cap="Cubble operations", out.height="50%", out.width="90%", fig.align='center'}
knitr::include_graphics(here::here("figures/cubble-operations/cubble-operations.001.png"))
```

### 2021-11-20 main.tex
change to JSS format

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

\title{A template for the \emph{arxiv} style}

\author{
    H.Sherry Zhang
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}} \\
   \And
    Dianne Cook
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}} \\
   \And
    Ursula Laa
   \\
    Institute of Statistics \\
    University of Natural Resources and Life Sciences \\
  Vienna, Austria \\
  \texttt{\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}} \\
   \And
    Nicolas Langrené
   \\
    34 Village Street, Docklands VIC 3008 Australia \\
    CSIRO Data61 \\
  Melbourne, Australia \\
  \texttt{\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}} \\
   \And
    Patricia Menéndez
   \\
    Department of Econometrics and Business Statistics \\
    Monash University \\
  Melbourne, Australia \\
  \texttt{\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}} \\
  }

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
Enter the text of your abstract here.
\end{abstract}

\keywords{
    blah
   \and
    blee
   \and
    bloo
   \and
    these are optional and can be removed
  }

@@

Many spatial and spatio-temporal data structures have been developed by
the R-spatial team for both raster and vector spatial data. For vector
spatial data, which is the focus of this paper, \texttt{sf} (E. J.
Pebesma 2018) represents spatial vector information with simple
features: points, lines, polygons and their multiples. Various
\texttt{st\_} function are designed to manipulate these features based
on their geometric relationships. For spatio-temporal data,
\texttt{stars} (E. Pebesma 2021) can represent both raster and vector
data using multi-dimensional array. However, the underlying array
structure can be difficult to operate for data analysts who are more
familiar with a flat 2D data frame structure used by the tidyverse
ecosystem.

In the temporal aspect, the \texttt{tsibble} (Wang, Cook, and Hyndman
2020) structure and its tidyverts ecosystem have provided a {[}\ldots{}
{]} workflow to work with temporal data. In a tsibble structure,
temporal data is characterised by \texttt{index} and \texttt{key} where
\texttt{index} is the temporal identifier and \texttt{key} is the
identifier for multiple series, which could be used as a spatio
identifier. However, a tsibble object, by construction, always requires
the \texttt{index} in its structure. This makes it less appealing for
spatio-temporal data since the output of calculated spatio-specific
variables (i.e.~features of each series) don't have the time dimension.
Analysts will either need to have an additional step to join this output
to the original tsibble or operate with variables stored in two separate
objects. In addition, the long form structure of a tsibble object means
spatio variables (i.e.~longitude, latitude, and features of each series
if joined back to the tsibble) of each spatio identifier will be
repetitively recorded at each timestamp. This repetition is unnecessary
and would inflate the object size for long series.

@@

\begin{figure}

@@

\caption{Cubble diagram}\label{fig:cubble-diagram}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{spatial\_new }\OtherTok{\textless{}{-}}\NormalTok{ spatial }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(SITES\_IN\_VICTORIA)}
\NormalTok{temporal\_new }\OtherTok{\textless{}{-}}\NormalTok{ temporal }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(id }\SpecialCharTok{\%in\%}\NormalTok{ spatial\_new}\SpecialCharTok{$}\NormalTok{id)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{climate\_flat }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: id-wise: nested form
## # Key:    id [5]
## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
##   id            lat  long  elev name           wmo_id ts                
##   <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
## 1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
## 2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
## 3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
## 4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
## 5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{verbatim}

@@

\begin{verbatim}
## # A tibble: 5 x 8
##   id            lat  long  elev name                 wmo_id  dist city     
##   <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>    
## 1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbourne
## 2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbourne
## 3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbourne
## 4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbourne
## 5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbourne
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sydmel\_climate }\OtherTok{\textless{}{-}}\NormalTok{ stations }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{rowwise}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{ts =} \FunctionTok{list}\NormalTok{(}\FunctionTok{meteo\_pull\_monitors}\NormalTok{(id, }
                                       \AttributeTok{date\_min =} \StringTok{"2020{-}01{-}01"}\NormalTok{, }
                                       \AttributeTok{date\_max =} \StringTok{"2020{-}12{-}31"}\NormalTok{,}
                                       \AttributeTok{var =} \FunctionTok{c}\NormalTok{(}\StringTok{"PRCP"}\NormalTok{, }\StringTok{"TMAX"}\NormalTok{, }\StringTok{"TMIN"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
                     \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{id))) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: id-wise: nested form
## # Key:    id [5]
## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
##   id            lat  long  elev name                 wmo_id  dist city   ts     
##   <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>  <list> 
## 1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbo~ <tibbl~
## 2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbo~ <tibbl~
## 3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbo~ <tibbl~
## 4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbo~ <tibbl~
## 5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbo~ <tibbl~
\end{verbatim}

@@

\begin{figure}

{\centering \includegraphics[width=1\linewidth,height=0.5\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-operations/cubble-operations.001}

@@

\caption{Cubble operations}\label{fig:cubble-operations}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{aus\_climate }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{==} \DecValTok{2020}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{tamp}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: id-wise: nested form
## # Key:    id [85]
## # Leaves: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
##    id            lat  long  elev name                   wmo_id ts               
##    <chr>       <dbl> <dbl> <dbl> <chr>                   <dbl> <list>           
##  1 ASN00009021 -31.9  116.  15.4 perth airport           94610 <tibble [366 x 4~
##  2 ASN00010311 -31.9  117. 179   york                    94623 <tibble [366 x 4~
##  3 ASN00010614 -32.9  117. 338   narrogin                94627 <tibble [366 x 4~
##  4 ASN00014015 -12.4  131.  30.4 darwin airport          94120 <tibble [366 x 4~
##  5 ASN00015131 -17.6  134. 220   elliott                 94236 <tibble [366 x 4~
##  6 ASN00016065 -30.4  137.  76   andamooka               95660 <tibble [366 x 4~
##  7 ASN00023034 -35.0  139.   2   adelaide airport        94672 <tibble [366 x 4~
##  8 ASN00023373 -34.5  139. 275   nuriootpa viticultural  94681 <tibble [366 x 4~
##  9 ASN00024024 -34.4  141.  30.1 loxton research centre  94682 <tibble [366 x 4~
## 10 ASN00024048 -34.2  141.  31.5 renmark aero            95687 <tibble [366 x 4~
## # ... with 75 more rows
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# join wmo\_id for each station}
\CommentTok{\# to\_join \textless{}{-} station \%\textgreater{}\% select(id, wmo\_id)}
\CommentTok{\# out \textless{}{-} climate\_small \%\textgreater{}\% }
\CommentTok{\#   left\_join(to\_join, by = c("station" = "id")) \%\textgreater{}\% }
\CommentTok{\#   filter(!is.na(wmo\_id))}
\CommentTok{\# out}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{aus\_climate }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{stretch}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{migrate}\NormalTok{(id, lat, long)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # Cubble: time-wise: long form
## # Key:    id [85]
## # Leaves: id [chr], lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id
## #   [dbl], prcp_missing [dbl], tmax_missing [dbl], tmin_missing [dbl]
##    id          date        prcp  tmax  tmin   lat  long
##    <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
##  1 ASN00009021 2020-01-01     0   319   153 -31.9  116.
##  2 ASN00009021 2020-01-02     0   249   164 -31.9  116.
##  3 ASN00009021 2020-01-03     6   232   130 -31.9  116.
##  4 ASN00009021 2020-01-04     0   284   124 -31.9  116.
##  5 ASN00009021 2020-01-05     0   353   116 -31.9  116.
##  6 ASN00009021 2020-01-06     0   348   131 -31.9  116.
##  7 ASN00009021 2020-01-07     0   328   151 -31.9  116.
##  8 ASN00009021 2020-01-08     0   304   174 -31.9  116.
##  9 ASN00009021 2020-01-09     0   287   173 -31.9  116.
## 10 ASN00009021 2020-01-10     0   326   158 -31.9  116.
## # ... with 31,100 more rows
\end{verbatim}

@@

\hypertarget{summary}{%
\section*{Summary}\label{summary}}
\addcontentsline{toc}{section}{Summary}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-stars}{}%
Pebesma, Edzer. 2021. \emph{Stars: Spatiotemporal Arrays, Raster and
Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.

\leavevmode\hypertarget{ref-pebesma2018simple}{}%
Pebesma, Edzer J. 2018. {``Simple Features for r: Standardized Support
for Spatial Vector Data.''} \emph{R J.} 10 (1): 439.

\leavevmode\hypertarget{ref-tsibbles}{}%
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``A New Tidy Data
Structure to Support Exploration and Modeling of Temporal Data.''}
\emph{Journal of Computational and Graphical Statistics} 29 (3):
466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\end{CSLReferences}

\bibliographystyle{unsrt}
\bibliography{references.bib}

>>>>

\documentclass[
]{jss}

\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash University
\AND Ursula Laa\\University of Natural Resources and Life Sciences
\AND Nicolas Langrené\\CSIRO Data61 \AND Patricia
Menéndez\\Monash University \AND
}
\title{\pkg{cubble}: An R Package for Structuring Spatio-temporal Data}

\Plainauthor{H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
Langrené, Patricia Menéndez}
\Plaintitle{cubble: An R Package for Structuring Spatio-temporal Data}

\Abstract{
The abstract of the article.
}

\Keywords{spatio-temporal data, \proglang{R}}
\Plainkeywords{spatio-temporal data, R}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
          }

% Pandoc citation processing

% Pandoc header

\usepackage{amsmath}

\begin{document}

@@

Many spatial and spatio-temporal data structures have been developed by
the R-spatial team for both raster and vector spatial data. For vector
spatial data, which is the focus of this paper, \texttt{sf}
\citep{pebesma2018simple} represents spatial vector information with
simple features: points, lines, polygons and their multiples. Various
\texttt{st\_} function are designed to manipulate these features based
on their geometric relationships. For spatio-temporal data,
\texttt{stars} \citep{stars} can represent both raster and vector data
using multi-dimensional array. However, the underlying array structure
can be difficult to operate for data analysts who are more familiar with
a flat 2D data frame structure used by the tidyverse ecosystem.

In the temporal aspect, the \texttt{tsibble} \citep{tsibbles} structure
and its tidyverts ecosystem have provided a {[}\ldots{} {]} workflow to
work with temporal data. In a tsibble structure, temporal data is
characterised by \texttt{index} and \texttt{key} where \texttt{index} is
the temporal identifier and \texttt{key} is the identifier for multiple
series, which could be used as a spatio identifier. However, a tsibble
object, by construction, always requires the \texttt{index} in its
structure. This makes it less appealing for spatio-temporal data since
the output of calculated spatio-specific variables (i.e.~features of
each series) don't have the time dimension. Analysts will either need to
have an additional step to join this output to the original tsibble or
operate with variables stored in two separate objects. In addition, the
long form structure of a tsibble object means spatio variables
(i.e.~longitude, latitude, and features of each series if joined back to
the tsibble) of each spatio identifier will be repetitively recorded at
each timestamp. This repetition is unnecessary and would inflate the
object size for long series.

@@

\begin{CodeChunk}
\begin{figure}

@@

\caption[Cubble diagram]{Cubble diagram}\label{fig:cubble-diagram}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> spatial_new <- spatial %>% filter(SITES_IN_VICTORIA)
R> temporal_new <- temporal %>% filter(id %in% spatial_new$id)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> climate_flat %>% as_cubble(key = id, index = date, coords = c(long, lat))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]- check gap on long and lat
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeOutput}
# A tibble: 5 x 8
  id            lat  long  elev name                 wmo_id  dist city     
  <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>    
1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbourne
2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbourne
3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbourne
4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbourne
5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbourne
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> sydmel_climate <- stations %>% 
+   rowwise() %>% 
+   mutate(ts = list(meteo_pull_monitors(id, 
+                                        date_min = "2020-01-01", 
+                                        date_max = "2020-12-31",
+                                        var = c("PRCP", "TMAX", "TMIN")) %>% 
+                      select(-id))) %>% 
+   as_cubble(key = id, index = date, coords = c(long, lat))
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [144.47, -38.03, 145.1, -37.38]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name                 wmo_id  dist city   ts     
  <chr>       <dbl> <dbl> <dbl> <chr>                 <dbl> <dbl> <chr>  <list> 
1 ASN00086038 -37.7  145.  78.4 essendon airport      95866  10.8 melbo~ <tibbl~
2 ASN00086282 -37.7  145. 113.  melbourne airport     94866  20.1 melbo~ <tibbl~
3 ASN00086077 -38.0  145.  12.1 moorabbin airport     94870  21.9 melbo~ <tibbl~
4 ASN00088162 -37.4  145. 528.  wallan (kilmore gap)  94860  48.1 melbo~ <tibbl~
5 ASN00087113 -38.0  144.  10.6 avalon airport        94854  48.8 melbo~ <tibbl~
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth,height=0.5\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/cubble-operations/cubble-operations.001}

@@

\caption[Cubble operations]{Cubble operations}\label{fig:cubble-operations}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> aus_climate %>% 
+   stretch() %>% 
+   filter(year(date) == 2020) %>% 
+   tamp()
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [85]: nested form
# bbox:     [115.97, -43.32, 153.13, -12.42]- check gap on long
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
   id            lat  long  elev name                   wmo_id ts               
   <chr>       <dbl> <dbl> <dbl> <chr>                   <dbl> <list>           
 1 ASN00009021 -31.9  116.  15.4 perth airport           94610 <tibble [366 x 4~
 2 ASN00010311 -31.9  117. 179   york                    94623 <tibble [366 x 4~
 3 ASN00010614 -32.9  117. 338   narrogin                94627 <tibble [366 x 4~
 4 ASN00014015 -12.4  131.  30.4 darwin airport          94120 <tibble [366 x 4~
 5 ASN00015131 -17.6  134. 220   elliott                 94236 <tibble [366 x 4~
 6 ASN00016065 -30.4  137.  76   andamooka               95660 <tibble [366 x 4~
 7 ASN00023034 -35.0  139.   2   adelaide airport        94672 <tibble [366 x 4~
 8 ASN00023373 -34.5  139. 275   nuriootpa viticultural  94681 <tibble [366 x 4~
 9 ASN00024024 -34.4  141.  30.1 loxton research centre  94682 <tibble [366 x 4~
10 ASN00024048 -34.2  141.  31.5 renmark aero            95687 <tibble [366 x 4~
# ... with 75 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # join wmo_id for each station
R> # to_join <- station %>% select(id, wmo_id)
R> # out <- climate_small %>% 
R> #   left_join(to_join, by = c("station" = "id")) %>% 
R> #   filter(!is.na(wmo_id))
R> # out
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> aus_climate %>% 
+   stretch() %>% 
+   migrate(id, lat, long)
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [85]: long form
# bbox:    [115.97, -43.32, 153.13, -12.42]- check gap on long
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
   id          date        prcp  tmax  tmin   lat  long
   <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
 1 ASN00009021 2020-01-01     0  31.9  15.3 -31.9  116.
 2 ASN00009021 2020-01-02     0  24.9  16.4 -31.9  116.
 3 ASN00009021 2020-01-03     6  23.2  13   -31.9  116.
 4 ASN00009021 2020-01-04     0  28.4  12.4 -31.9  116.
 5 ASN00009021 2020-01-05     0  35.3  11.6 -31.9  116.
 6 ASN00009021 2020-01-06     0  34.8  13.1 -31.9  116.
 7 ASN00009021 2020-01-07     0  32.8  15.1 -31.9  116.
 8 ASN00009021 2020-01-08     0  30.4  17.4 -31.9  116.
 9 ASN00009021 2020-01-09     0  28.7  17.3 -31.9  116.
10 ASN00009021 2020-01-10     0  32.6  15.8 -31.9  116.
# ... with 31,100 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\renewcommand\refname{Summary}
\bibliography{references.bib}
